### 2015-05-13 whoop.tex
addressed some comments

Symbolic verification of a pair of entry points works by (i) instrumenting each entry point with additional state to record locksets, and (ii) attempting to verify a sequential program that executes the instrumented entry points in sequence and then asserts, for each shared location, that the locksets for each entry point with respect to this location have a non-empty intersection.  Verification of the resulting sequential program can be undertaken using any sound method; in practice we employ the Boogie verification engine~\cite{barnett2006boogie}, which requires procedure specifications and loop invariants to be generated, after which verification conditions are generated and discharged to an automated theorem prover.

@@

\medskip\noindent\textbf{Verification and Error Reporting }
%
For each pair of entry points, the instrumented sequential program, equipped with procedure and loop summaries, is sent to the Boogie verification engine. Boogie generates a verification condition~\cite{barnett2005weakest}, for each procedure in the program, and discharges it to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair. Successful verification implies that the entry point pair is free of data races, while an error (i.e.\ counterexample) denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is a race that \whoop found and reported for the example of Figure~\ref{fig:data_race_example}:

@@

First, we enriched \whoop with information regarding \emph{kernel-imposed serialization}, to increase precision. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise.  \ADComment{Can you make it clearer that this information is currently provided manually to \whoop?  Can you explain how this is achieved?  And can you comment briefly on how we figured this out?}

%Modeling kernel methods is another opportunity to increase precision, but requires Linux kernel expertise. For example, the \texttt{register\_netdev()} function registers a network driver with the kernel. Before this function is called, the kernel is not able to invoke any of the driver's entry points. Modeling this functionality can reduce false positives when race-checking accesses before the driver's registration.

Second, we soundly reduce the number of memory regions that are analyzed for races.  If memory region $\mathit{MR}$ is accessed by only one entry point in a pair then, trivially, the pair cannot race on $\mathit{MR}$.  We thus disable lockset analysis for $\mathit{MR}$.  This can reduce the complexity of verification conditions that need to be solved by the theorem prover, speeding up the verification process.

>>>>

Symbolic verification of a pair of entry points works by (i) instrumenting each entry point with additional state to record locksets, and (ii) attempting to verify a sequential program that executes the instrumented entry points in sequence and then asserts, for each shared location, that the locksets for each entry point with respect to this location have a non-empty intersection.  Verification of the resulting sequential program can be undertaken using any sound method; in practice we employ the Boogie verification engine~\cite{barnett2006boogie}, which requires procedure specifications and loop invariants to be generated, after which verification conditions~\cite{barnett2005weakest} (VCs) are generated and discharged to an automated theorem prover.

@@

\medskip\noindent\textbf{Verification and Error Reporting }
%
For each pair of entry points, the instrumented sequential program, equipped with procedure and loop summaries, is sent to the Boogie verification engine. Boogie generates a VC, for each procedure in the program, and discharges it to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair. Successful verification implies that the entry point pair is free of data races, while an error (i.e.\ counterexample) denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is a race that \whoop found and reported for the example of Figure~\ref{fig:data_race_example}:

@@

First, we enriched \whoop with information regarding \emph{kernel-imposed serialization}, to increase precision. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. We discovered this when we first analyzed the r8169 driver (see \S\ref{evaluation}) and \whoop reported many races between a number of networking entry points; when we investigated the source of these races, we found out that these entry points could not race in reality because of RTNL. \whoop exploits this knowledge and does not create pairs for entry points that are serialized by the kernel. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise. Currently, this information is hard-coded inside \whoop. In the future, we plan to provide a configuration file where the user can denote such kernel specific information.

Second, we soundly reduce the number of memory regions that are analyzed for races.  If memory region $\mathit{MR}$ is accessed by only one entry point in a pair then, trivially, the pair cannot race on $\mathit{MR}$.  We thus disable lockset analysis for $\mathit{MR}$.  This can reduce the complexity of VCs that need to be solved by the theorem prover, speeding up the verification process.
